---
title: "Test Adaptive Algorithms for Educational Games"
author: "Gloria and George"
date: "7/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```


## Define the Stimuli and Difficulties

Our pilot will use upper- and lower-case letters, ordered by decreasing frequency (here ordered alphabetically for convenience).

```{r define-stimuli}
literacy_stim = c(LETTERS, letters)
numeracy_stim = c(1:25, 0, 26:51)

# priority queues (priority scale is arbitrary, but higher = easier)
litQ = data.frame(stim=literacy_stim, priority=seq(length(literacy_stim), 1))
numQ = data.frame(stim=numeracy_stim, priority=seq(length(numeracy_stim), 1))
```


## Define Adjustment Function

The adaptive algorithm works in the following way:

```{r pressure, echo=FALSE}
CHUNK_SIZE = 8
CHUNK_PRIORITY_INCREASE = .5
CORRECT_PRIORITY_INCREASE = 1
INCORRECT_PRIORITY_INCREASE = .5
CORRECT_PROB_INCREASE = .5 #.05
INCORRECT_PROB_INCREASE = .05 # .01
NUM_TRIALS = 520 
# realistic estimate 560: 70 trials/5 mins, exposure ~40 mins
# chose 520 based on completing standard curriculum 2x
MIN_FOILS = 2
MAX_FOILS = 9
num_foils = 2
STANDARD_NUM_FOILS = 5
```


## Define Learners

Let's start with learners who have a fixed probability *r* of correctly responding to each stimulus.

## Standard Curriculum (SC) Simulations

```{r simulate-standard-subject}

# try half .7 and half .4
standard_learner_ability = 1

# could model learning if we increase these probs (say, if they get it correct + .1)
lit_prob_correct = rep(standard_learner_ability, length(literacy_stim))

sigmoid <- function(x) { 1/(1+exp(-x)) }

simulate_standard_subject <- function(stimuli, prob_correct, num_trials = 100, learning = FALSE) {
  standard_literacy_stim = rep(stimuli, each = 5, times = 2) # 520 trials
  #standard_literacy_stim = c(standard_literacy_stim, standard_literacy_stim[1:40])
  
  # track data we want to plot
  stim_dat <- tibble(stimulus = stimuli, 
                     prob_correct = prob_correct,
                     freq = 0, # how many times each stim occurs
                     correct = 0) # if we model learning, track this
            
  # pick next easiest, simulate correct or not, update
  prev_correct = F
  for (t in seq(num_trials)) {
      cur_stim = standard_literacy_stim[t]
      cur_stim_idx = which(stim_dat$stimulus==cur_stim)
        stim_dat[cur_stim_idx,]$freq = stim_dat[cur_stim_idx,]$freq + 1 # increase frequency count
    
    p = stim_dat[cur_stim_idx,]$prob_correct / STANDARD_NUM_FOILS
    if(p>1) p = 1 # can't be >1
    correcto = rbinom(1, 1, p)
    
    if (correcto) {
      if (learning == TRUE) {
        stim_dat[cur_stim_idx,]$prob_correct =
          stim_dat[cur_stim_idx,]$prob_correct + CORRECT_PROB_INCREASE
      }
      stim_dat[cur_stim_idx,]$correct = stim_dat[cur_stim_idx,]$correct + 1 
      prev_correct = T
    } else { # wrong
      # decrease difficulty
      if (learning == TRUE) {
        stim_dat[cur_stim_idx,]$prob_correct =
          stim_dat[cur_stim_idx,]$prob_correct + INCORRECT_PROB_INCREASE
      }
      prev_correct = F
    }
  }
  return(stim_dat)
}

simulate_standard_subject(literacy_stim, .5, num_trials = 520, learning = TRUE)

```

```{r simulate-standard-experiment}
simulate_standard_experiment <- function(stimuli, prob_correct, num_trials, learning, Nsubj=10) {
  dat <- tibble()
  for(i in 1:Nsubj) {
    s_dat = simulate_standard_subject(stimuli, prob_correct, num_trials, learning)
    s_dat$sim = i
    dat <- bind_rows(dat, s_dat)
  }
  return(dat)
}

# test 
e_dat <- simulate_standard_experiment(literacy_stim, lit_prob_correct, 
                 num_trials = NUM_TRIALS, learning = F, Nsubj=30)

# summarise results 
agg <- e_dat %>% group_by(stimulus) %>%
  mutate(prop_correct = correct / freq) %>%
  summarise(freq = mean(freq), prop_correct = mean(prop_correct)) 

  # summarise(freq = mean(freq), priority_delta = mean(priority_delta),
  #           prop_correct_Q1 = prop_correct = mean(prop_correct),
  #           prop_correct_Q2 = prop_correct = mean(prop_correct),
  #           prop_correct_Q3 = prop_correct = mean(prop_correct),
  #           prop_correct_Q4 = prop_correct = mean(prop_correct)) 
# View(agg)
agg$stimulus <- ordered(agg$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
agg %>% ggplot(aes(y=stimulus, x=prop_correct)) + 
  geom_point() + theme_bw() + xlim(0, 1)

ggsave(file="SC-1-F.png", width=5, height=5)
# file name: "CONDITION-ABILITY-LEARNING"
```

## Adaptive Algorithm (AA) Simulations

```{r simulate-adaptive-subject}

# 1 for chance=1/num_foils; 2 for 2U1D; 3.5 for solid learner
learner_ability = 3.5

# could model learning if we increase these
# probs (say, if they get it correct + .1)
lit_prob_correct = rep(learner_ability, length(literacy_stim)) 
num_prob_correct = rep(learner_ability, length(numeracy_stim)) 

simulate_adaptive_subject <- function(stimuli, prob_correct, num_trials = 100, learning = FALSE, adaptive = FALSE) {

  # track data we want to plot
  stim_dat <- tibble(stimulus = stimuli, 
                     prob_correct = prob_correct,
                     freq = 0, # how many times each stim occurs
                     correct = 0, # if we model learning, track this
                     start_priority = 0, 
                     cur_priority = 0)
  
  # initialize priorities
  pr = 0
  for (i in 1:nrow(stim_dat)) {
    if (i%%CHUNK_SIZE==0) pr = pr + CHUNK_PRIORITY_INCREASE
    pr = pr + 0.01
    stim_dat[i,]$start_priority = pr
    stim_dat[i,]$cur_priority = pr
  }
  
  # pick next easiest, simulate correct or not, update
  prev_correct = F
  for (t in seq(num_trials)) {
    curr_index = which(stim_dat$cur_priority==min(stim_dat$cur_priority)) # find index
    stim_dat[curr_index,]$freq = stim_dat[curr_index,]$freq + 1 # increase frequency count
    
    p = stim_dat[curr_index,]$prob_correct / num_foils 
    if(p>1) p = 1 # can't be >1
    correcto = rbinom(1, 1, p)
    
    if (correcto) { # && frequency <= 5
      stim_dat[curr_index,]$cur_priority = 
        stim_dat[curr_index,]$cur_priority + CORRECT_PRIORITY_INCREASE
      stim_dat[curr_index,]$correct = stim_dat[curr_index,]$correct + 1 # increase correct count
      if(prev_correct & t > 1 & num_foils < MAX_FOILS) num_foils = num_foils + 1
      if (learning == TRUE) {
        stim_dat[curr_index,]$prob_correct =
          stim_dat[curr_index,]$prob_correct + CORRECT_PROB_INCREASE
      }
      prev_correct = T
    } else { # wrong
      # decrease difficulty
      if(num_foils > MIN_FOILS) num_foils = num_foils - 1
      stim_dat[curr_index,]$cur_priority = 
        stim_dat[curr_index,]$cur_priority + INCORRECT_PRIORITY_INCREASE
      if (learning == TRUE) {
        stim_dat[curr_index,]$prob_correct =
          stim_dat[curr_index,]$prob_correct + INCORRECT_PROB_INCREASE
      }
      prev_correct = F
    }
  }
  return(stim_dat)
}

#simulate_adaptive_subject(literacy_stim, .5, num_trials = 520, learning = TRUE)
```

```{r simulate-adaptive-experiment}
simulate_adaptive_experiment <- function(stimuli, prob_correct, num_trials, learning, Nsubj=10) {
  dat <- tibble()
  for(i in 1:Nsubj) {
    s_dat = simulate_adaptive_subject(stimuli, prob_correct, num_trials, learning)
    s_dat$sim = i
    dat <- bind_rows(dat, s_dat)
  }
  return(dat)
}

# test 
e_dat <- simulate_adaptive_experiment(literacy_stim, lit_prob_correct, 
                 num_trials = NUM_TRIALS, learning = T, Nsubj=30)

# summarise results 
agg <- e_dat %>% group_by(stimulus) %>%
  mutate(prop_correct = correct / freq,
         priority_delta = cur_priority - start_priority) %>%
  summarise(freq = mean(freq), priority_delta = mean(priority_delta),
            prop_correct = mean(prop_correct)) 

  # summarise(freq = mean(freq), priority_delta = mean(priority_delta),
  #           prop_correct_Q1 = prop_correct = mean(prop_correct),
  #           prop_correct_Q2 = prop_correct = mean(prop_correct),
  #           prop_correct_Q3 = prop_correct = mean(prop_correct),
  #           prop_correct_Q4 = prop_correct = mean(prop_correct)) 
# View(agg)
agg$stimulus <- ordered(agg$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
agg %>% ggplot(aes(y=stimulus, x=prop_correct)) + 
  geom_point() + theme_bw() + xlim(0, 1)

ggsave(file="AA-3.5-T.png", width=5, height=5)
# file name: "CONDITION-ABILITY-LEARNING"
```


``` {r make-tables}

make_new_correct_table <- function(orig_table) {
  new_table <- tibble(stimulus = orig_table$stimulus, 
                      answer_type = "correct",
                      frequency = orig_table$correct)
  return(new_table)
}

make_new_incorrect_table <- function(orig_table) {
  new_table <- tibble(stimulus = orig_table$stimulus, 
                      answer_type = " incorrect",
                      frequency = orig_table$freq - orig_table$correct)
  return(new_table)
}

make_new_new_table <- function(new_table) {
  new_new_table <- add_row(make_new_incorrect_table(new_table), make_new_correct_table(new_table)) 
  return(new_new_table)
}

```

## Now a couple of simulations

``` {r simulation-1}
# 1) baseline .5 random model, with no learning

lit_sim_pt5 = simulate_subject(literacy_stim, lit_prob_correct, num_trials = NUM_TRIALS, learning = FALSE)
# View(lit_sim_pt5)

e_dat_lit_sim_pt5 <- simulate_experiment(literacy_stim, lit_prob_correct, 
                 num_trials = NUM_TRIALS, learning = F, Nsubj=30)

View(e_dat_lit_sim_pt5)
new_e_dat_lit_sim_pt5_table <- make_new_new_table(e_dat_lit_sim_pt5)
View(new_e_dat_lit_sim_pt5_table)
# summarise results 
agg <- new_e_dat_lit_sim_pt5_table %>% group_by(stimulus) %>%
  summarise(freq = mean(frequency)) 

View(agg)
new_lit_sim_pt5_table$stimulus <- ordered(new_lit_sim_pt5_table$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
# View(new_lit_sim_pt5_table)         

lit_sim_pt5_plot <- ggplot(data=new_lit_sim_pt5_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
lit_sim_pt5_plot

num_sim_pt5 = simulate_subject(numeracy_stim, num_prob_correct, num_trials = NUM_TRIALS, learning = FALSE)
# View(num_sim_pt5)
new_num_sim_pt5_table <- make_new_new_table(num_sim_pt5)
num_sim_pt5_plot <- ggplot(data=new_num_sim_pt5_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
num_sim_pt5_plot
```

``` {r simulation-2}
# 2) "realistic": better knowledge of easier stimuli

lit_prob_correct = 1 - seq(.01, 1, 1/length(literacy_stim))
lit_sim_easier <- simulate_subject(literacy_stim, lit_prob_correct, num_trials = NUM_TRIALS, learning = FALSE)
# View(lit_sim_easier)
new_lit_sim_easier_table <- make_new_new_table(lit_sim_easier)
# View(new_lit_sim_easier_table)  
new_lit_sim_easier_table$stimulus <- ordered(new_lit_sim_easier_table$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
 #View(new_lit_sim_easier_table)                                         

lit_sim_easier_plot <- ggplot(data=new_lit_sim_easier_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
lit_sim_easier_plot
```

``` {r simulation-3}
# 3) some kind of learning... increase prob_correct as function of freq  (up to 1)
lit_prob_correct = 1 - seq(.01, 1, 1/length(literacy_stim))
lit_sim_learning <- simulate_subject(literacy_stim, lit_prob_correct, num_trials = NUM_TRIALS, learning = TRUE)
# View(lit_sim_learning)
new_lit_sim_learning_table <- make_new_new_table(lit_sim_learning)
# View(new_lit_sim_learning_table)  
new_lit_sim_learning_table$stimulus <- ordered(new_lit_sim_learning_table$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
 #View(new_lit_sim_learning_table)                                         

lit_sim_learning_plot <- ggplot(data=new_lit_sim_learning_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
lit_sim_learning_plot

# plot freq, change in priority 


```

``` {r sim-plots}
library(ggplot2)
library(RColorBrewer)
library(ggthemes)

```

