---
title: "Test Adaptive Algorithms for Educational Games"
author: "George"
date: "7/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
```


## Define the Stimuli and Difficulties

Our pilot will use upper- and lower-case letters, ordered by decreasing frequency (here ordered alphabetically for convenience).

```{r define-stimuli}
literacy_stim = c(LETTERS, letters)
numeracy_stim = c(1:25, 0, 26:51)

# priority queues (priority scale is arbitrary, but higher = easier)
litQ = data.frame(stim=literacy_stim, priority=seq(length(literacy_stim), 1))
numQ = data.frame(stim=numeracy_stim, priority=seq(length(numeracy_stim), 1))
```


## Define Adjustment Function

The adaptive algorithm works in the following way:

```{r pressure, echo=FALSE}
CHUNK_SIZE = 8
CHUNK_PRIORITY_INCREASE = .5
CORRECT_PRIORITY_INCREASE = 1
INCORRECT_PRIORITY_INCREASE = .5
CORRECT_PROB_INCREASE = .1 #.05
INCORRECT_PROB_INCREASE = .05 # .01
NUM_TRIALS = 70 * 8 # realistic estimate: 70 trials/5 mins, exposure ~40 mins

```


## Define Learners

Let's start with learners who have a fixed probability *r* of correctly responding to each stimulus.

```{r example-pq}

# 0.5 for chance; 0.66 for 2U1D; 0.95 for solid learner
# could model learning if we increase these
lit_prob_correct = rep(.5, length(literacy_stim)) 
num_prob_correct = rep(.5, length(numeracy_stim)) 
# probs (say, if they get it correct + .1)

run_exp <- function(stimuli, prob_correct, num_trials = 100, learning = FALSE) {
  # consider adding NUM_FOILS that changes as we play--changes the probability of correct choice
  
  # track data we want to plot
  stim_dat <- tibble(stimulus = stimuli, 
                     prob_correct = prob_correct,
                     freq = 0, # how many times each stim occurs
                     correct = 0, # if we model learning, track this
                     start_priority = 0, 
                     cur_priority = 0)
  
  # initialize priorities
  pr = 0
  for (i in 1:nrow(stim_dat)) {
    if (i%%CHUNK_SIZE==0) pr = pr + CHUNK_PRIORITY_INCREASE
    pr = pr + 0.01
    stim_dat[i,]$start_priority = pr
    stim_dat[i,]$cur_priority = pr
  }
  
  # pick next easiest, simulate correct or not, update
  for (t in seq(num_trials)) {
    curr_index = which(stim_dat$cur_priority==min(stim_dat$cur_priority)) # find index
    stim_dat[curr_index,]$freq = stim_dat[curr_index,]$freq + 1 # increase frequency count
    randnum <- runif(1, min=0, max=1) 
    # consider replacing with rbinom and using NFOILS to change probability of success
    
    if (randnum < stim_dat[curr_index,]$prob_correct) {
      stim_dat[curr_index,]$cur_priority = 
        stim_dat[curr_index,]$cur_priority + CORRECT_PRIORITY_INCREASE
      stim_dat[curr_index,]$correct = stim_dat[curr_index,]$correct + 1 # increase correct count
      if (learning == TRUE && stim_dat[curr_index,]$prob_correct + CORRECT_PROB_INCREASE < 1) {
        stim_dat[curr_index,]$prob_correct =
          stim_dat[curr_index,]$prob_correct + CORRECT_PROB_INCREASE
      }
    } else {
      stim_dat[curr_index,]$cur_priority = 
        stim_dat[curr_index,]$cur_priority + INCORRECT_PRIORITY_INCREASE
      if (learning == TRUE && stim_dat[curr_index,]$prob_correct + INCORRECT_PROB_INCREASE < 1) {
        stim_dat[curr_index,]$prob_correct =
          stim_dat[curr_index,]$prob_correct + INCORRECT_PROB_INCREASE
      }
    }
  }
  return(stim_dat)
}
```

``` {r make-tables}

make_new_correct_table <- function(orig_table) {
  new_table <- tibble(stimulus = orig_table$stimulus, 
                      answer_type = "correct",
                      frequency = orig_table$correct)
  return(new_table)
}

make_new_incorrect_table <- function(orig_table) {
  new_table <- tibble(stimulus = orig_table$stimulus, 
                      answer_type = " incorrect",
                      frequency = orig_table$freq - orig_table$correct)
  return(new_table)
}

make_new_new_table <- function(new_table) {
  new_new_table <- add_row(make_new_incorrect_table(new_table), make_new_correct_table(new_table)) 
  return(new_new_table)
}

```

## Now a couple of simulations

``` {r stimulation-1}
# 1) baseline .5 random model, with no learning

lit_sim_pt5 = run_exp(literacy_stim, lit_prob_correct, num_trials = NUM_TRIALS, learning = FALSE)
# View(lit_sim_pt5)
new_lit_sim_pt5_table <- make_new_new_table(lit_sim_pt5)

new_lit_sim_pt5_table$stimulus <- ordered(new_lit_sim_pt5_table$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
# View(new_lit_sim_pt5_table)         

lit_sim_pt5_plot <- ggplot(data=new_lit_sim_pt5_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
lit_sim_pt5_plot

num_sim_pt5 = run_exp(numeracy_stim, num_prob_correct, num_trials = NUM_TRIALS, learning = FALSE)
# View(num_sim_pt5)
new_num_sim_pt5_table <- make_new_new_table(num_sim_pt5)
num_sim_pt5_plot <- ggplot(data=new_num_sim_pt5_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
num_sim_pt5_plot
```

``` {r simulation-2}
# 2) "realistic": better knowledge of easier stimuli

lit_prob_correct = 1 - seq(.01, 1, 1/length(literacy_stim))
lit_sim_easier <- run_exp(literacy_stim, lit_prob_correct, num_trials = NUM_TRIALS, learning = FALSE)
# View(lit_sim_easier)
new_lit_sim_easier_table <- make_new_new_table(lit_sim_easier)
# View(new_lit_sim_easier_table)  
new_lit_sim_easier_table$stimulus <- ordered(new_lit_sim_easier_table$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
 #View(new_lit_sim_easier_table)                                         

lit_sim_easier_plot <- ggplot(data=new_lit_sim_easier_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
lit_sim_easier_plot
```

``` {r simulation-3}
# 3) some kind of learning... increase prob_correct as function of freq  (up to 1)
lit_prob_correct = 1 - seq(.01, 1, 1/length(literacy_stim))
lit_sim_learning <- run_exp(literacy_stim, lit_prob_correct, num_trials = NUM_TRIALS, learning = TRUE)
# View(lit_sim_learning)
new_lit_sim_learning_table <- make_new_new_table(lit_sim_learning)
# View(new_lit_sim_learning_table)  
new_lit_sim_learning_table$stimulus <- ordered(new_lit_sim_learning_table$stimulus, levels=c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N", "O", "P", "Q", "R", "S", "T", "U", "V", "W", "X", "Y", "Z", "a", "b", "c", "d", "e", "f", "g", "h", "i", "j", "k", "l", "m", "n", "o", "p", "q", "r", "s", "t", "u", "v", "w", "x", "y","z"))
 #View(new_lit_sim_learning_table)                                         

lit_sim_learning_plot <- ggplot(data=new_lit_sim_learning_table, aes(x=stimulus, y=frequency, fill=answer_type)) + geom_bar(stat="identity")
lit_sim_learning_plot

# plot freq, change in priority 


```

``` {r sim-plots}
library(ggplot2)
library(RColorBrewer)
library(ggthemes)

```

